---
title: "Exercise_3_Xiru Lyu"
author: "Xiru Lyu"
date: "1/26/2018"
output: pdf_document
---

## Problem 1

### (a) Write functions $\texttt{tmpFn1}$ and $\texttt{tmpFn2}$ such that if $\texttt{xVec}$ is the vector $(x_{1},x_{2},...,x_{n})$, then $\texttt{tmpFn1(xVec)}$ returns the vector $(x_{1},x_{2}^{2},...,x_{n}^{n})$ and $\texttt{tmpFn2(xVec)}$ returns the vector $(x_{1},\frac{x_{2}^{2}}{2},...,\frac{x_{n}^{n}}{n})$.

```{r}
tmpFn1 <- function(xVec) {
  for (i in 1:length(xVec)) {
    xVec[i] <- xVec[i]^i
  }
  xVec
}

tmpFn2 <- function(xVec) {
  for (i in 1:length(xVec)){
    xVec[i] <- xVec[i]^i/i}
  xVec
}
```

### (b) Now write a function $\texttt{tmpFn3}$ which takes 2 arguments $\texttt{x}$ and $\texttt{n}$ where $\texttt{x}$ is a single number and $\texttt{n}$ is a strictly positive integer. The function should return the value of $$1+\frac{x}{1}+\frac{x^{2}}{2}+\frac{x^{3}}{3}+...+\frac{x^{n}}{n}$$
```{r}
tmpFn2 <- function(x,n) {
sum(x^(1:n)/(1:n)) + 1
}
```


## Problem 2

### Write a function $\texttt{tmpFn(xVec)}$ such that if $\texttt{xVec}$ is the vector $x = (x_{1},x_{2},...,x_{n})$ then $\texttt{tmpFn(xVec)}$ returns the vector of moving averages: $$\frac{x_{1}+x_{2}+x_{3}}{3},\frac{x_{2}+x_{3}+x_{4}}{3},...,\frac{x_{n-2}+x_{n-1}+x_{n}}{3}$$
```{r}
tmpFn <- function(xVec) {
  # create a new vector
  new <- numeric()
  
  for (i in 3:length(xVec)) {
    new <- c(new,(xVec[i-2]+xVec[i-1]+xVec[i])/3)
  }
  new
}
```


## Problem 3

### Consider the continuous function 
$$f(x)=\left\{\begin{matrix}
 x^{2}+2x+3 & \texttt{if } x<0 \\ 
 x+3 &  \texttt{if } 0\leq x<2\\ 
 x^{2}+4x-7&   \texttt{if } 2\leq x
\end{matrix}\right.$$

### Write a function $\texttt{tmpFn}$ which takes a single argument $\texttt{xVec}$. The function should return the vector of values of the function *f(x)* evaluated at the values in $\texttt{xVec}$.
```{r}
tmpFn <- function(xVec){
  if (xVec < 0){
    xVec^2+2*xVec+3
  } else if (xVec < 2) {
    xVec+3
  } else {
    xVec^2+4*xVec-7
  }
}
```

### Hence plot the function *f(x)* for $-3<x<3$.
```{r warning=FALSE}
# set up the range of the plot
range <- seq(-3,3,length=1000)

# make the plot
plot(range,tmpFn(range),type='l')
```


## Problem 4

### Write a function which takes a single argument which is a matrix. The function should return a matrix which is the same as the function argument but every odd number is doubled. Hence the result of using the function on the matrix 
$$\begin{bmatrix}
1 & 1 & 3 \\ 
5 & 2 & 6 \\ 
-2 & -1  & -3 
\end{bmatrix} $$

### should be 
$$\begin{bmatrix}
2 & 2 & 6 \\ 
10 & 2 & 6 \\ 
-2 & -2  & -6 
\end{bmatrix} $$

```{r}
fun1 <- function(matrix) {
  matrix[matrix%%2==1] <- 2*matrix[matrix%%2==1]
  matrix
}
```


## Problem 5

### Write a function which takes 2 arguments $\texttt{n}$ and $\texttt{k}$ which are positive integers. It should return the *n × n* matrix:
$$\begin{bmatrix}
k & 1 & 0 & 0 & ... & 0 & 0 \\ 
1 & k & 1 & 0 & ... & 0 & 0 \\ 
0 & 1 & k & 1 & ... & 0 & 0 \\ 
0 & 0 & 1 & k & ... & 0 & 0 \\ 
... & ... & ... & ... & ... & ... & ... \\
0 & 0 & 0 & 0 & ... & k & 1 \\ 
0 & 0 & 0 & 0 & ... & 1 & k
\end{bmatrix} $$
```{r}
fun2 <- function(n,k) {
    new <- diag(k,ncol=n,nrow=n)
    new[abs(col(new)-row(new))==1] <- 1
    new
}
```


## Problem 6

### Suppose an angle $\alpha$ is given as a positive real number of degrees.

### If $0\leq\alpha<90$ then it is quadrant 1. If $90\leq\alpha<180$ then it is quadrant 2. If $180\leq\alpha<270$ then it is quadrant 3. If $270\leq\alpha<360$ then it is quadrant 4. If $360\leq\alpha<450$ then it is quadrant 1. And so on.

### Write a function $\texttt{quadrant(alpha)}$ which returns the quadrant of the angle $\alpha$.

```{r}
quadrant <- function(alpha) {
  if (alpha < 360) {
    if (alpha >= 0 & alpha < 90) {
      print('quadrant 1')
    } else if (alpha < 180) {
      print('quadrant 2')
    } else if (alpha < 270) {
      print('quadrant 3')
    } else {
      print('qudrant 4')
    }
  }
  else {
    alpha_new <- alpha - 360
    if (alpha_new >= 0 & alpha_new < 90){
      print('quadrant 1')
    } else if (alpha_new < 180) {
      print('quadrant 2')
    } else if (alpha_new < 270) {
      print('qudrant 3')
    } else {
      print('quadrant 4')
    }
  }
}
```


## Problem 7

### Zeller’s congruence is the formula:
$$f=([2.6m-0.2]+k+y+[y/4]+[c/4]-2c)\mod7$$

### where [*x*] denotes the integer part of *x*

### Write a function $\texttt{weekday(day,month,year)}$ which returns the day of the week when given the nu- merical inputs of the day, month and year.
```{r}
weekday <- function(day,month,year) {
  k <- day
  y <- year %% 100
  c <- trunc(year / 100)
  if (month < 3) {
    m <- month + 10
    y <- y - 1
  } else {
    m <- month - 2
  }
  f <- (trunc(2.6*m-0.2) + k + y + trunc(y/4) + trunc(c/4) - 2*c) %% 7 + 1
  c('Sunday','Monday','Tuesday','Wednesday','Thursday','Friday','Saturday')[f]
}
```

### (b) Does your function work if the input parameters $\texttt{day}$, $\texttt{month}$ and $\texttt{year}$ are vectors with the same length and with valid entries?
```{r}
# the function needs to have a part to check the validity of each entry

weekday2 <- function(day,month,year) {
  if (month < 0 | month > 12) {
    break
  } else if (day < 0 | day > 31) {
    break
  } else {
    k <- day
    y <- year %% 100
    c <- trunc(year / 100)
    if (month < 3) {
      m <- month + 10
      y <- y - 1
    } else {
    m <- month - 2
    }
    f <- (trunc(2.6*m-0.2) + k + y + trunc(y/4) + trunc(c/4) - 2*c) %% 7 + 1
    c('Sunday','Monday','Tuesday','Wednesday','Thursday','Friday','Saturday')[f]
  }
}
```

## Problem 8

### (a) Suppose $x_{0}=1$ and $x_{1}=2$ and 
$$x_{j}=x_{j-1}+\frac{2}{x_{j-1}}\quad\textrm{and}\;j=1,2,....$$

### Write a function $\texttt{testLoop}$ which takes the single argument *n* and returns the first *n* $-\;1$ values of the sequence {x$_{j}$}$_{j\geq0}$: that means the values of $x_{0},x_{1},x_{2},...,x_{n-2}$.
```{r}
testLoop <- function(n) {
  x <- rep(NA,n-1)
  x[1] <- 1
  x[2] <- 2
  if (n-1==2) {
    x
  }
  else {
  for (i in 3:n-1) 
    {x[i] <- x[i-1] + 2/x[i-1]} 
  x }
  }
```

### (b) Now write a funtoin $\texttt{testLoop2}$ which takes a single argument $\texttt{yVec}$ which is a vector. The function should return
$$\sum_{j=1}^{n}e^{j}$$

### where *n* is the length of $\texttt{yVec}$
```{r}
testLoop2 <- function(yVec) {
  n <- length(yVec)
  seq <- 1:n
  sum(exp(seq))
}
```


## Problem 9

### (a) Write a function $\texttt{quadmap(start,rho,niter)}$ which returns the vector ($x_{1},...,x_{n}$) where $x_{k}=rx_{k-1}(1-x_{k-1})$
```{r}
quadmap <- function(start,rho,niter) {
  x <- rep(NA,niter)
  x[1] <- start
  for (i in 2:niter) {
    x[i] <- rho*x[i-1]*(1-x[i-1])
  }
  x
}
```

### Try out the function you have written
```{r}
tmp <- quadmap(start=0.95, rho=2.99, niter=500)

# try the plot
plot(tmp,type='l')

# try another plot
plot(tmp[300:500], type="l")
```

### (b) Now write a function which determines the number of iterations needed to get $\left | x_{n}-x_{n-1} \right |<0.02$. So this function only has two arguments.
```{r}
quadmap2 <- function(start,rho){
  x_1 <- start
  niter <- 1
  x_2 <- rho*x_1*(1-x_1)
  while (abs(x_2-x_1) >= 0.02) {
    x_1 <- x_2
    x_2 <- rho*x_1*(1-x_1)
    niter <- niter + 1
  }
  niter
}

# test the function
# the function should return 84 for start=0.95, rho=2.99
quadmap2(start=0.95,rho=2.99)
```


## Problem 10

### (a) Given a vector ($x_{1},...,x_{n}$), the sample autocorrelation $\texttt{k}$ is defined to be $$r_{k}=\frac{\sum_{i=k+1}^{n}(x_{i}-\bar{x})(x_{i-k}-\bar{x})}{\sum_{i=1}^{n}(x_{i}-\bar{x})^{2}}$$Thus$$r_{1}=\frac{\sum_{i=2}^{n}(x_{i}-\bar{x})(x_{i-1}-\bar{x})}{\sum_{i=1}^{n}(x_{i}-\bar{x})^{2}}=\frac{(x_{2}-\bar{x})(x_{1}-\bar{x})+...+(x_{n}-\bar{x})(x_{n-1}-\bar{x})}{\sum_{i=1}^{n}(x_{i}-\bar{x})^{2}}$$Write a function $\texttt{tmpFn(xVec)}$ which takes a single argument $\texttt{xVec}$ and returns a $\texttt{list}$ of two values:$r_{1}$ and $r_{2}$.

### In particular, find $r_{1}$ and $r_{2}$ for the vector (2, 5, 8, . . . , 53, 56).
```{r}
tmpFn <- function(xVec) {
  
  # calculate the value of r1
  i <- 1:length(xVec)
  i_1 <- 2:length(xVec)
  num_1 <- sum((xVec[i_1]-mean(xVec))*(xVec[i_1-1]-mean(xVec))) # find the numerator
  denom <- sum((xVec[i]-mean(xVec))^2) # find the denominator
  r1 <- num_1/denom # compute r1
  
  # calculate the value of r2
  i_2 <- 3:length(xVec)
  num_2 <- sum((xVec[i_2]-mean(xVec))*(xVec[i_2-2]-mean(xVec))) # find the numerator
  r2 <- num_2/denom # compute r2
  
  # Return the list with r1 and r2
  r <- list(r1 = r1,r2 = r2)
  r
}

# find r1 and r2 for the vector
vec <- seq(from=2,to=56,by=3)
tmpFn(vec)
```


### (b) Generalise the function so that it takes two arguments: the vector $\texttt{xVec}$ and an integer $\texttt{k}$ which lines between 1 and *n* $-\;1$ where *n* is the length of $\texttt{xVec}$.

### The function should return a vector of the values ($r_{0} = 1,r_{1},...,r_{k}$).
```{r}
tmpFn_2 <- function(xVec,k) {
  # Write a function that generates the value for each k
  tmp <- function(k) {
    xbar <- mean(xVec)
  
    # Calculate the value of the denominator
    length_d <- seq(from=1, to=length(xVec))
    denom <- sum((xVec[length_d]-xbar)^2)
    
    # Calculate the value of the numerator
    length_n <- seq(from=k+1, to=length(xVec))
    num_1 <- sum((xVec[length_n]-xbar)*(xVec[length_n-k]-xbar))
  
    r <- num_1/denom }
  
  list <- c(1,sapply(1:k,tmp))
  list
}
```

