---
title: "Exercise_2_Xiru Lyu"
author: "Xiru Lyu"
date: "1/25/2018"
output:
  pdf_document: default
---

## Problem 1

### Suppose 
$$ A = \begin{bmatrix}
1 & 1 & 3 \\ 
5 & 2 & 6 \\ 
-2 & -1  & -3 
\end{bmatrix} $$

### (a) Check that $A^{3}= 0$ where 0 is a 3×3 matrix with every entry equal to 0.
```{r}
A <- cbind(c(1,5,-2),c(1,2,-1),c(3,6,-3))
A %*% A %*% A
```

### (b) Replace the third column of A by the sum of the second and third columns.
```{r}
x <- A[,2] + A[,3]
A[,3] <- x
A
```

## Problem 2

### Create the following matrix B with 15 rows: 
$$\begin{bmatrix}
10 & -10 & 10 \\ 
10 & -10 & 10 \\
... & ... & ... \\
10 & -10  & 10 
\end{bmatrix}$$

### Calculate the 3 × 3 matrix **B**$^{T}$**B**.

```{r}
B <- matrix(nrow=15,ncol=3)
B[,1] <- 10; B[,2] <- -10; B[,3] <- 10
t(B) %*% B
```

## Problem 3

### Create a 6 × 6 matrix $\texttt{matE}$ with every entry equal to 0. Check what the functions row and col return when applied to $\texttt{matE}$. Hence create the 6 × 6 matrix:
$$\begin{bmatrix}
0 & 1 & 0 & 0 & 0 & 0 \\ 
1 & 0 & 1 & 0 & 0 & 0 \\
0 & 1 & 0 & 1 & 0 & 0 \\
0 & 0 & 1 & 0 & 1 & 0 \\
0 & 0 & 0 & 1 & 0 & 1 \\
0 & 0 & 0 & 0 & 1 & 0 
\end{bmatrix}$$

```{r}
matE <- matrix(0,nrow=6,ncol=6)
matE[abs(col(matE)-row(matE)) == 1] <- 1
matE
```


## Problem 4

### Create the following patterned matrix:
$$\begin{bmatrix}
0 & 1 & 2 & 3 & 4 \\ 
1 & 2 & 3 & 4 & 5 \\
2 & 3 & 4 & 5 & 6 \\
3 & 4 & 5 & 6 & 7 \\
4 & 5 & 6 & 7 & 8
\end{bmatrix}$$

```{r}
outer(0:4,0:4,"+")
```


## Problem 5: Create the following patterned matrices

### (a) 
$$\begin{bmatrix}
0 & 1 & 2 & 3 & 4 \\ 
1 & 2 & 3 & 4 & 0 \\
2 & 3 & 4 & 0 & 1 \\
3 & 4 & 0 & 1 & 2 \\
4 & 0 & 1 & 2 & 3
\end{bmatrix}$$
```{r}
outer(0:4,0:4,"+") %% 5
```

### (b)
$$\begin{bmatrix}
0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 \\ 
1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 0 \\ 
. & . & . & . & . & . & . & . & . & . \\ 
. & . & . & . & . & . & . & . & . & . \\ 
. & . & . & . & . & . & . & . & . & . \\ 
8 & 9 & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7  \\ 
9 & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8  
\end{bmatrix}$$

```{r}
outer(0:9,0:9,"+") %% 10
```

### (c)
$$\begin{bmatrix}
0 & 8 & 7 & 6 & 5 & 4 & 3 & 2 & 1 \\ 
1 & 0 & 8 & 7 & 6 & 5 & 4 & 3 & 2 \\ 
2 & 1 & 0 & 8 & 7 & 6 & 5 & 4 & 3 \\ 
3 & 2 & 1 & 0 & 8 & 7 & 6 & 5 & 4 \\ 
4 & 3 & 2 & 1 & 0 & 8 & 7 & 6 & 5 \\ 
5 & 4 & 3 & 2 & 1 & 0 & 8 & 7 & 6 \\
6 & 5 & 4 & 3 & 2 & 1 & 0 & 8 & 7 \\
7 & 6 & 5 & 4 & 3 & 2 & 1 & 0 & 8 \\
8 & 7 & 6 & 5 & 4 & 3 & 2 & 1 & 0 
\end{bmatrix}$$

```{r}
e <- outer(0:8,0:8,"+")
f <- row(e) - col (e)
for (i in 1:9){
  for (j in 1:9){
    if (f[i,j] < 0){
      f[i,j] <- f[i,j] + nrow(f)
    }}}
f
```


## Problem 6: Solve the System
```{r}
m <- matrix(0,nrow=5,ncol=5)
a <- abs(col(m)-row(m))+1
b <- rbind(7,-1,-3,5,17)
solve(a,b)
```

## Problem 7

```{r}
set.seed(75)
aMat <- matrix( sample(10, size=60, replace=T), nr=6)
```

### (a) Find the number of entries in each row which are greater than 4.
```{r}
apply(aMat,1,function(x){sum(x>4)})
```

### (b) Which rows contain exactly two occurrences of the number seven?
```{r}
which(apply(aMat,1,function(x){sum(x==7)==2}))
```

### (c) Find those pairs of columns whose total (over both columns) is greater than 75.
```{r}
# Find sums for each column
aMat_colsum <- apply(aMat,2,sum)

# Find the outer product for the sum
outer <- outer(aMat_colsum,aMat_colsum,"+")

# Exclude cases when the value of the outer product is computed by the sum of the same column 
for (i in 1:nrow(outer)) {
  for (j in 1:ncol(outer)) {
    if (i == j) {
      outer[i,j] <- 0
    }
  }
}

# Find col numbers whose total is greater than 75
which(outer>75,arr.ind=TRUE)
```


## Problem 8

### (a) Calculate $\sum_{i=1}^{20}\sum_{j=1}^{5}\frac{i^4}{(3+j)}$
```{r}
# Create a matrix that contains all possible values with i varies between 1 
# and 20, j varies between 1 and 5
mat1 <- matrix(nrow=20,ncol=5)
for (i in 1:nrow(mat1)){
  for (j in 1:ncol(mat1)){
    mat1[i,j] <- i^4/(3+j)
  }
}

# Find the sum of all these values in the matrix mat1
sum(mat1)
```

### (b) Calculate $\sum_{i=1}^{20}\sum_{j=1}^{5}\frac{i^4}{(3+ij)}$
```{r}
# Create a matrix that contains all possible values with i varies between 1 
# and 20, j varies between 1 and 5
mat2 <- matrix(nrow=20,ncol=5)
for (i in 1:nrow(mat2)) {
  for (j in 1:ncol(mat2)) {
    mat2[i,j] <- i^4/(3+i*j)
  }
}

# Find the sum of all these values in the matrix mat2
sum(mat2)
```

### (c) Calculate $\sum_{i=1}^{10}\sum_{j=1}^{i}\frac{i^4}{(3+ij)}$
```{r}
# Create a matrix that contains all possible values with i varies between 1 
# and 10, j varies between 1 and 10 (i)
mat3 <- matrix(0,nrow=10,ncol=10)
for (i in 1:nrow(mat3)) {
  for (j in 1:i) {
    mat3[i,j] <- i^4/(3+i*j)
  }
}

# Find the sum all of these values in the matrix mat3
sum(mat3)
```

