---
title: "Exercise_3_Xiru Lyu"
author: "Xiru Lyu"
date: "1/26/2018"
output: pdf_document
---

## Problem 1

### (a) Write functions $\texttt{tmpFn1}$ and $\texttt{tmpFn2}$ such that if $\texttt{xVec}$ is the vector $(x_{1},x_{2},...,x_{n})$, then $\texttt{tmpFn1(xVec)}$ returns the vector $(x_{1},x_{2}^{2},...,x_{n}^{n})$ and $\texttt{tmpFn2(xVec)}$ returns the vector $(x_{1},\frac{x_{2}^{2}}{2},...,\frac{x_{n}^{n}}{n})$.

```{r}
tmpFn1 <- function(xVec) {
  for (i in 1:length(xVec)) {
    xVec[i] <- xVec[i]^2
  }
  xVec
}
```

### (b) Now write a function $\texttt{tmpFn3}$ which takes 2 arguments $\texttt{x}$ and $\texttt{n}$ where $\texttt{x}$ is a single number and $\texttt{n}$ is a strictly positive integer. The function should return the value of $$1+\frac{x}{1}+\frac{x^{2}}{2}+\frac{x^{3}}{3}+...+\frac{x^{n}}{n}$$
```{r}
tmpFn2 <- function(xVec) {
  for (i in 1:length(xVec)) {
    xVec[i] <- xVec[i]^2/i
  }
  xVec
}
```


## Problem 2

### Write a function $\texttt{tmpFn(xVec)}$ such that if $\texttt{xVec}$ is the vector $x = (x_{1},x_{2},...,x_{n})$ then $\texttt{tmpFn(xVec)}$ returns the vector of moving averages: $$\frac{x_{1}+x_{2}+x_{3}}{3},\frac{x_{2}+x_{3}+x_{4}}{3},...,\frac{x_{n-2}+x_{n-1}+x_{n}}{3}$$
```{r}
tmpFn <- function(xVec) {
  # create a new vector
  new <- numeric()
  
  for (i in 3:length(xVec)) {
    new <- c(new,(xVec[i-2]+xVec[i-1]+xVec[i])/3)
  }
  new
}
```


## Problem 3

### Consider the continuous function 
$$f(x)=\left\{\begin{matrix}
 x^{2}+2x+3 & \texttt{if } x<0 \\ 
 x+3 &  \texttt{if } 0\leq x<2\\ 
 x^{2}+4x-7&   \texttt{if } 2\leq x
\end{matrix}\right.$$

### Write a function $\texttt{tmpFn}$ which takes a single argument $\texttt{xVec}$. The function should return the vector of values of the function *f(x)* evaluated at the values in $\texttt{xVec}$.
```{r}
tmpFn <- function(xVec){
  if (xVec < 0){
    xVec^2+2*xVec+3
  } else if (xVec < 2) {
    xVec+3
  } else {
    xVec^2+xVec-7
  }
}
```

### Hence plot the function *f(x)* for $-3<x<3$.
```{r warning=FALSE}
# set up the range of the plot
range <- seq(-3,3,length=1000)

# make the plot
plot(range,tmpFn(range),type='l')
```


## Problem 4

### Write a function which takes a single argument which is a matrix. The function should return a matrix which is the same as the function argument but every odd number is doubled. Hence the result of using the function on the matrix 
$$\begin{bmatrix}
1 & 1 & 3 \\ 
5 & 2 & 6 \\ 
-2 & -1  & -3 
\end{bmatrix} $$

### should be 
$$\begin{bmatrix}
2 & 2 & 6 \\ 
10 & 2 & 6 \\ 
-2 & -2  & -6 
\end{bmatrix} $$

```{r}
fun1 <- function(matrix) {
  # create a new matrix that is a copy of the input matrix
  new_matrix <- matrix
  
  for (i in 1:nrow(matrix)) {
    for (j in 1:ncol(matrix)) {
      if (matrix[i,j] %% 2 == 1) {
        new_matrix[i,j] <- 2*new_matrix[i,j]
      }
    }
  }
  new_matrix
}
```


## Problem 5

### Write a function which takes 2 arguments $\texttt{n}$ and $\texttt{k}$ which are positive integers. It should return the *n × n* matrix:
$$\begin{bmatrix}
k & 1 & 0 & 0 & ... & 0 & 0 \\ 
1 & k & 1 & 0 & ... & 0 & 0 \\ 
0 & 1 & k & 1 & ... & 0 & 0 \\ 
0 & 0 & 1 & k & ... & 0 & 0 \\ 
... & ... & ... & ... & ... & ... & ... \\
0 & 0 & 0 & 0 & ... & k & 1 \\ 
0 & 0 & 0 & 0 & ... & 1 & k
\end{bmatrix} $$
```{r}
fun2 <- function(n,k) {
    new <- diag(k,ncol=n,nrow=n)
    new[abs(col(new)-row(new))==1] <- 1
    new
}
```


## Problem 6

### Suppose an angle $\alpha$ is given as a positive real number of degrees.

### If $0\leq\alpha<90$ then it is quadrant 1. If $90\leq\alpha<180$ then it is quadrant 2. If $180\leq\alpha<270$ then it is quadrant 3. If $270\leq\alpha<360$ then it is quadrant 4. If $360\leq\alpha<450$ then it is quadrant 1. And so on.

### Write a function $\texttt{quadrant(alpha)}$ which returns the quadrant of the angle $\alpha$.

```{r}
quadrant <- function(alpha) {
  if (alpha < 360) {
    if (alpha >= 0 & alpha < 90) {
      print('quadrant 1')
    } else if (alpha < 180) {
      print('quadrant 2')
    } else if (alpha < 270) {
      print('quadrant 3')
    } else {
      print('qudrant 4')
    }
  }
  else {
    alpha_new <- alpha - 360
    if (alpha_new >= 0 & alpha_new < 90){
      print('quadrant 1')
    } else if (alpha_new < 180) {
      print('quadrant 2')
    } else if (alpha_new < 270) {
      print('qudrant 3')
    } else {
      print('quadrant 4')
    }
  }
}
```


## Problem 7

### Zeller’s congruence is the formula:
$$f=([2.6m-0.2]+k+y+[y/4]+[c/4]-2c)\mod7$$

### where [*x*] denotes the integer part of *x*

### Write a function $\texttt{weekday(day,month,year)}$ which returns the day of the week when given the nu- merical inputs of the day, month and year.
```{r}
weekday <- function(day,month,year) {
  k <- day
  y <- year %% 100
  c <- trunc(year / 100)
  if (month < 3) {
    m <- month + 10
    y <- y - 1
  } else {
    m <- month - 2
  }
  f <- (trunc(2.6*m-0.2) + k + y + trunc(y/4) + trunc(c/4) - 2*c) %% 7 + 1
  c('Sunday','Monday','Tuesday','Wednesday','Thursday','Friday','Saturday')[f]
}
```

### (b) Does your function work if the input parameters $\texttt{day}$, $\texttt{month}$ and $\texttt{year}$ are vectors with the same length and with valid entries?
```{r}
# the function needs to have a part to check the validity of each entry

weekday2 <- function(day,month,year) {
  if (month < 0 | month > 12) {
    break
  } else if (day < 0 | day > 31) {
    break
  } else {
    k <- day
    y <- year %% 100
    c <- trunc(year / 100)
    if (month < 3) {
      m <- month + 10
      y <- y - 1
    } else {
    m <- month - 2
    }
    f <- (trunc(2.6*m-0.2) + k + y + trunc(y/4) + trunc(c/4) - 2*c) %% 7 + 1
    c('Sunday','Monday','Tuesday','Wednesday','Thursday','Friday','Saturday')[f]
  }
}
```

## Problem 8

### (a) Suppose $x_{0}=1$ and $x_{1}=2$ and 
$$x_{j}=x_{j-1}+\frac{2}{x_{j-1}}\quad\textrm{and}\;j=1,2,....$$

### Write a function $\texttt{testLoop}$ which takes the single argument *n* and returns the first *n* $-\;1$ values of the sequence {x$_{j}$}$_{j\geq0}$: that means the values of $x_{0},x_{1},x_{2},...,x_{n-2}$.
```{r}

```

### (b) Now write a funtoin $\texttt{testLoop2}$ which takes a single argument $\texttt{yVec}$ which is a vector. The function should return
$$\sum_{j=1}^{n}e^{j}$$

### where *n* is the length of $\texttt{yVec}$
```{r}

```


## Problem 9

### (a) Write a function $\texttt{quadmap(start,rho,niter)}$ which returns the vector ($x_{1},...,x_{n}$) where $x_{k}=rx_{k-1}(1-x_{k-1})$
```{r}

```

### Try out the function you have written
```{r}

```

### Try the plot
```{r}

```

### Try another plot
```{r}

```

### (b) Now write a function which determines the number of iterations needed to get $\left | x_{n}-x_{n-1} \right |<0.02$. So this function only has two arguments.


## Problem 10

### (a) Given a vector ($x_{1},...,x_{n}$), the sample autocorrelation $\texttt{k}$ is defined to be $$r_{k}=\frac{\sum_{i=k+1}^{n}(x_{i}-\bar{x})(x_{i-k}-\bar{x})}{\sum_{i=1}^{n}(x_{i}-\bar{x})^{2}}$$Thus$$r_{1}=\frac{\sum_{i=2}^{n}(x_{i}-\bar{x})(x_{i-1}-\bar{x})}{\sum_{i=1}^{n}(x_{i}-\bar{x})^{2}}=\frac{(x_{2}-\bar{x})(x_{1}-\bar{x})+...+(x_{n}-\bar{x})(x_{n-1}-\bar{x})}{\sum_{i=1}^{n}(x_{i}-\bar{x})^{2}}$$Write a function $\texttt{tmpFn(xVec)}$ which takes a single argument $\texttt{xVec}$ and returns a $\texttt{list}$ of two values:$r_{1}$ and $r_{2}$.

### In particular, find $r_{1}$ and $r_{2}$ for the vector (2, 5, 8, . . . , 53, 56).
```{r}

```


### (b) Generalise the function so that it takes two arguments: the vector $\texttt{xVec}$ and an integer $\texttt{k}$ which lines between 1 and *n* $-\;1$ where *n* is the length of $\texttt{xVec}$.

### The function should return a vector of the values ($r_{0} = 1,r_{1},...,r_{k}$).
